# Data Visualization

In this section, you will learn:

1. The structure of `ggplot`.
2. A few techniques to make your plots better.
3. A taste of animation.

In this section, we will use the following libraries and data:

```{r}
library(tidyverse)
library(gganimate)

slgname <-
  read_csv("data_sources/saved_data/slgname.csv")
```

To drive towards an end goal as we learn about `ggplot`, lets create a visualization to depict the annual hit totals for players with the top 20 total career hits.

## Bar Chart

```{r, cache = TRUE}
slgname
```

### Canvas {-}

The base layer of any ggplot object is simply the "canvas" in which you will build your plot.

```{r, cache = TRUE}
slgname %>% 
  ggplot()
```

### Mapping {-}

We will "map" each element of the data to the canvas.

Elements which can be mapped include `x`, `y`, `color`, `fill`, `size`, `alpha`, and a few others.

```{r, cache = TRUE}
slgname %>% 
  ggplot(aes(x=player, y = hr))
```

### Filter Data {-}

To keep the plot clean, we'll filter for the top 20 players by career hits.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>% 
  # top_n(20, h) %>% 
  ggplot(aes(x=player, y = hr)) 
```

### Plot 'mechanism' {-}

Next we add the type of plot.  There are a *ton* and can be explored [here](https://ggplot2.tidyverse.org/reference/).  

We will only explore a couple.

Also, we should point out the simplicity of ggplot.  Each 'layer' is added sequentially.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>%  
  ggplot(aes(x=player, y = hr)) +
  geom_col()
```

### Coordinate Flip {-}

Bar plots often suffer from difficult to read x axes.  We can fix that with a coordinate flip.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>%  
  ggplot(aes(x=player, y = hr)) +
  geom_col() +
  coord_flip()
```

### Reorder Factors {-}

This is nice, but it doesn't easily present the information we are trying to convey.  Lets reorder the factors of the barplot.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>%  
  ggplot(aes(x=fct_reorder(player,hr), y = hr)) +
  geom_col() +
  coord_flip()
```

### Color Bars {-}

Lets add some more information to the plot by coloring the bars by the players slugging percentage.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>%  
  ggplot(aes(x=fct_reorder(player,hr), y = hr, fill = slg)) +
  geom_col() +
  coord_flip()
```

### Update Labels {-}

Lets update our plot labels to help the plot tell the story of the data.

Notice in the `labs()` function, you simply provide the information you would like to present to the mapped value.

```{r, cache = TRUE}
slgname %>% 
  slice_max(order_by = h, n = 20) %>%  
  ggplot(aes(x=fct_reorder(player,hr), y = hr, fill = slg)) +
  geom_col() +
  coord_flip() +
  labs(x = "Player", y = "Home Runs", 
       title = "Top 20 Home Run Hitters", 
       fill = "Slugging Percentage",
       subtitle = "1871-2016", 
       caption = "*Among Players Who've Played at least 1000 games")
```

## Scatter Plot

Next, lets show the functionality of the scatter plot.  

Lets explore the trajectory of the home runs per season since 1990

Lets start with our original data:

```{r, cache = TRUE}
data <- read_csv("data_sources/Batting.csv", col_types = cols(SF = col_double(), GIDP = col_double())) %>% 
  clean_names()

data
```


```{r, cache = TRUE}
scatterdat <-
  data %>% 
  filter(lg_id == "AL") %>% 
  filter(year_id >= 1990) %>% 
  mutate(team_id = fct_lump(team_id, n = 12)) %>%
  filter(team_id != "Other") %>% 
  filter(complete.cases(.)) %>%
  group_by(team_id,year_id) %>% 
  summarise(across(.cols = c(g:gidp),.fns = sum)) ## dplyr 1.0.0


scatterdat
```

### Canvas {-}

```{r, cache = TRUE}
scatterdat %>% 
  ggplot() 
```

### Mapping {-}

```{r, cache = TRUE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr,color = team_id))
```

### Plot 'mechanism': geom_point() {-}

`geom_point()` creates a scatterplot.

```{r, cache = TRUE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr,color = team_id)) +
  geom_point() 
```

### Arrange Teams by Least to Most Home Runs

Using `fct_reorder`, we can order the teams by most to least home runs.

```{r, cache = TRUE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr,color = fct_reorder(team_id,-hr))) +
  geom_point() 
```

### Plot 'mechanism': geom_smooth() {-}

We can stack layers to help tell the story of our data.

```{r, cache = TRUE, warning = FALSE, message = FALSE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth()
```

How does `ggplot` determine that line?

LOWESS - LOcally WEighted Scatterplot Smoothing

LOESS - LOcally Estimated Scatterplot Smoothing

Controlled by the term `span`.  Smaller equals more "wigglyness"

`span = 1`

```{r, cache = TRUE, warning = FALSE, message = FALSE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth(span = 1)
```

`span = .1`

```{r, cache = TRUE, warning = FALSE, message = FALSE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth(span = .1)
```

Both

```{r, cache = TRUE, warning = FALSE, message = FALSE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth(span = .1) +
  geom_smooth(span = 1, color = "red")
```

### Facet Plots {-}

`ggplot` offers the ability to 'facet' plots by a variable.  This can help show contrast between different factors.

```{r, cache = TRUE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth() +
  facet_wrap(~team_id)
```

### Update Labels {-}

As before, we can update the labels.

```{r, cache = TRUE, warning = FALSE, message = FALSE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth() +
  facet_wrap(~team_id) +
  labs(title = "Home Run Changes Over Time",
       subtitle = str_c("From ", min(scatterdat$year_id), " to ", max(scatterdat$year_id)),
       color = "Team",
       x = "Season", y = "Season Home Run Total")
```

### Change Smoothing Line {-}

Earlier, we talked about the LOESS smoother.  We can change the line type we impose on our scatter plots.  In this instance, we can apply a best fit linear regression line. 

```{r, cache = TRUE}
scatterdat %>% 
  ggplot(aes(x=year_id,y=hr)) +
  geom_point(aes(color = fct_reorder(team_id,-hr))) +
  geom_smooth(method = "lm") +
  facet_wrap(~team_id) +
  labs(title = "Home Run Changes Over Time",
       subtitle = paste("From", min(scatterdat$year_id), "to", max(scatterdat$year_id)),
       color = "Team",
       x = "Season", y = "Season Home Run Total")
```

## Animation

While not rendered here, we will briefly display the ability to animate a ggplot.  We've created some new data to support animation showing the change in the number of home runs and strike outs over the years.

```{r animate_data, echo=TRUE, message=FALSE, warning=FALSE, cache = TRUE}
animdata <-
  data %>%
  filter(!is.na(lg_id)) %>%
  group_by(year_id, team_id) %>%
  summarise(across(.cols = c(hr, so, bb, ab), .fns =  ~ sum(., na.rm = TRUE))) %>%
  ungroup() %>%
  inner_join(data %>%
               select(year_id, team_id, lg_id) %>%
               distinct()
             ) 

animdata
```

We'll talk through the code below.  While it does not render in the book, we'll run it in console and talk through the process.

There are many gganimate options and you can explore further [here](https://gganimate.com/articles/gganimate.html)

```{r animate_run, eval=FALSE, cache=TRUE}
p <- animdata %>% 
  ggplot(aes(x = so,y = hr, color = bb,size = ab,group = year_id)) +
  geom_point() +
  facet_wrap(~ lg_id) +
  transition_states(year_id,transition_length = 1,state_length = 30) +
  labs(title = "The Change in Home Runs and Strike Outs Over The Years", subtitle = 'Year: {closest_state}', x = "Strike Outs", y = "Home Runs") +
  # labs(title = 'Year: {closest_state, cache = TRUE}') +
  enter_fade() +
  exit_fade()

animate(p,nframes = length(unique(data$year_id)) * 2)

```



<!-- ```{r, cache = TRUE} -->
<!-- library(gapminder) -->

<!-- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop, colour = country)) + -->
<!--   geom_point(alpha = 0.7, show.legend = FALSE) + -->
<!--   scale_colour_manual(values = country_colors) + -->
<!--   scale_size(range = c(2, 12)) + -->
<!--   scale_x_log10() + -->
<!--   facet_wrap(~continent) + -->
<!--   # Here comes the gganimate specific bits -->
<!--   labs(title = 'Year: {frame_time, cache = TRUE}', x = 'GDP per capita', y = 'life expectancy') + -->
<!--   transition_time(year) + -->
<!--   ease_aes('linear') -->

<!-- # file_renderer(dir = ".", prefix = "gganim_plot", overwrite = FALSE) -->
<!-- ``` -->

