# Data Modeling

## Load Packages

```{r}
library(tidymodels)
library(tune)
library(dials)
library(parsnip)
library(rsample)
library(recipes)
library(textrecipes)
library(yardstick)
library(vip)
library(gghighlight)
library(patchwork)
library(tidyverse)
library(tidyquant)
library(knitr)
library(janitor)
```

## Read Data

```{r}
data = read_csv("Batting.csv", col_types = cols(SF = col_double(), GIDP = col_double())) %>%
  clean_names()

hofdata = read_csv("HallOfFame.csv") %>% 
  clean_names() %>% 
  select(player_id, inducted) %>% 
  mutate(hof = ifelse(inducted=="Y",1,0)) %>% 
  filter(hof==1)

hofdata %>%  count(inducted)

retiredyear = 
data %>% 
  group_by(player_id) %>% 
  summarise(lastyear = max(year_id))

hof =
data %>% 
  group_by(player_id) %>% 
  summarise_at(vars(g:gidp), list(~sum(.,na.rm = TRUE))) %>%
  ungroup() %>%  
  left_join(retiredyear) %>% 
  left_join(hofdata) %>% 
  mutate(inducted = if_else(inducted=="Y",1,0)) %>% 
  mutate(inducted = replace_na(inducted, 0)) %>% 
  filter(g>=1000) %>% 
  mutate(inducted = as.factor(inducted)) %>% 
  select(-hof)

hof

hof %>% 
  filter(lastyear <= 2012) %>% 
  count(inducted) 

hofmod =
hof %>% 
  filter(lastyear <= 2012) %>% 
  select(-lastyear)
hofmod

hoftest = 
hof %>% 
  filter(lastyear > 2012) %>% 
  select(-lastyear,-inducted)
hoftest

```

## Data Exploration

```{r}
hofmod %>% 
  select(g:inducted) %>%  
  pivot_longer(cols = g:gidp) %>% 
  ggplot(aes(x=value,y=as.factor(inducted),color = name)) +
  geom_point() +
  theme(legend.position = "none") +
  facet_wrap(~name, scales = "free")
```

## Split Data test/train

```{r}
set.seed(as.numeric(as.factor("beatnavy")))
hof_initial_split <- initial_split(hofmod, prop = 0.80)
hof_initial_split
```

## Preprocess Data

```{r}
preprocessing_recipe <-
  recipe(inducted ~ ., data = training(hof_initial_split)) %>%
  step_knnimpute(all_numeric()) %>% 
  step_center(all_numeric()) %>% 
  step_scale(all_numeric()) %>% 
  step_rm(player_id) %>%
  prep()
preprocessing_recipe
```

## Apply Preprocessing

```{r}
hof_training_preprocessed_tbl = 
  preprocessing_recipe %>% 
  bake(training(hof_initial_split))

hof_training_preprocessed_tbl
```

## Prepare Cross Validation

```{r}
set.seed(as.numeric(as.factor("beatnavy")))
hof_cv_folds =
  training(hof_initial_split) %>% 
    bake(preprocessing_recipe, new_data = .) %>%
    vfold_cv(v = 5)
hof_cv_folds
```

## Specify Models 

### GLM Model {-}

```{r}
glmnet_model =
  logistic_reg(mode = "classification",
    penalty = tune(),
    mixture = tune()) %>%
  set_engine("glmnet")

glmnet_model
```

### Random Forest Model {-}

```{r}
forest_model =
  rand_forest(
    mode = "classification",
    mtry = tune(),
    trees = tune(),
    min_n = tune()
  ) %>%
  set_engine("randomForest", objective = "reg:squarederror")

forest_model
```

## Create Grid of Parameters to Validate Over

### GLM Model {-}

```{r}
glmnet_params = parameters(penalty(), mixture())
glmnet_params

set.seed(as.numeric(as.factor("beatnavy")))
glmnet_grid <- grid_max_entropy(glmnet_params, size = 20)
glmnet_grid

glmnet_grid %>%
  ggplot(aes(penalty, mixture)) +
  geom_point(size = 3) +
  scale_x_log10() +
  labs(title = "Max Entropy Grid", x = "Penalty (log scale)", y = "Mixture")
```

### Random Forest Model {-}

```{r}
forest_params <- parameters(mtry(c(2,6)), trees(), min_n())
forest_params

set.seed(as.numeric(as.factor("beatnavy")))
forest_grid <- grid_max_entropy(forest_params, size = 30)
forest_grid

```

## Execute Cross Validation

### GLM Model {-}

```{r message=FALSE, warning=FALSE}
library(tune)
glmnet_stage_1_cv_results_tbl <- tune_grid(
    formula   = inducted ~ .,
    model     = glmnet_model,
    resamples = hof_cv_folds,
    grid      = glmnet_grid,
    metrics   = metric_set(accuracy, kap, roc_auc),
    control   = control_grid(verbose = TRUE)
)

glmnet_stage_1_cv_results_tbl %>% show_best("accuracy", n = 10, maximize = FALSE)
glmnet_stage_1_cv_results_tbl %>% show_best("kap", n = 10, maximize = FALSE)
glmnet_stage_1_cv_results_tbl %>% show_best("roc_auc", n = 10, maximize = FALSE)
```

### Random Forest Model {-}

```{r message=FALSE, warning=FALSE}
forest_stage_1_cv_results_tbl <- tune_grid(
    formula   = inducted ~ .,
    model     = forest_model,
    resamples = hof_cv_folds,
    grid      = forest_grid,
    metrics   = metric_set(accuracy, kap, roc_auc),
    control   = control_grid(verbose = TRUE)
)

forest_stage_1_cv_results_tbl %>% show_best("accuracy", n = 10, maximize = FALSE)
forest_stage_1_cv_results_tbl %>% show_best("kap", n = 10, maximize = FALSE)
forest_stage_1_cv_results_tbl %>% show_best("roc_auc", n = 10, maximize = FALSE)

```

## Select Best Parameters

```{r}
params_glmnet_best <- glmnet_stage_1_cv_results_tbl %>% 
    select_best("roc_auc", maximize = FALSE)
params_glmnet_best

params_forest_best <- forest_stage_1_cv_results_tbl %>% 
    select_best("roc_auc", maximize = FALSE)
params_forest_best
```

## Save Best Paramenters

```{r}
glmnet_stage_2_model <- glmnet_model %>% 
    finalize_model(parameters = params_glmnet_best)

glmnet_stage_2_model
```

```{r}
forest_stage_2_model <- forest_model %>% 
    finalize_model(params_forest_best)

forest_stage_2_model
```

## Compare Models

```{r}
train_processed <- 
  training(hof_initial_split) %>% 
  bake(preprocessing_recipe, new_data = .)

test_processed  <- 
  testing(hof_initial_split) %>% 
  bake(preprocessing_recipe, new_data = .)
    
target_expr <- preprocessing_recipe %>% 
    pluck("last_term_info") %>%
    filter(role == "outcome") %>%
    pull(variable) %>%
    sym()

glmnet_stage_2_metrics =    
  glmnet_stage_2_model %>%
  fit(formula = inducted ~ ., data = train_processed) %>%
  predict(new_data = test_processed) %>%
  bind_cols(testing(hof_initial_split)) %>%
  metrics(!! target_expr, .pred_class)

forest_stage_2_metrics =    
  forest_stage_2_model %>%
  fit(formula = inducted ~ ., data = train_processed) %>%
  predict(new_data = test_processed) %>%
  bind_cols(testing(hof_initial_split)) %>%
  metrics(!! target_expr, .pred_class)

glmnet_stage_2_metrics %>% 
  mutate(mod = "glmnet") %>% 
  bind_rows(
    forest_stage_2_metrics %>% 
      mutate(mod = "forest")
    ) %>% 
  arrange(.metric,-.estimate)

```

Looks like the Random Forest is the better model.

## Run Best Model on All Data

```{r}
model_final <- forest_stage_2_model %>%
    fit(inducted ~ . , data = bake(preprocessing_recipe, new_data = hofmod))
```

## Run Model on New Data

```{r}
hoftest %>%
  bake(preprocessing_recipe, new_data = .) %>%
  predict(model_final, new_data = .) %>% 
  bind_cols(hoftest %>% select(player_id)) %>% 
  arrange(desc(.pred_class))
```

## Variable Importance

```{r}
vip(model_final) +
    labs(title = "Random Forest Model Importance - HOF Prediction") 
```


[Many Thanks](https://www.r-bloggers.com/product-price-prediction-a-tidy-hyperparameter-tuning-and-cross-validation-tutorial/)

